---
title: "Final Exam"
output: html_document
---

```{r, echo=FALSE, message=FALSE}
rm(list=ls())
```

```{r, echo=FALSE, message=FALSE}
library(caret)
library(tidyverse)
library(dplyr)
library(ISLR)
library(factoextra)
library(corrplot)
library(fpc)
```

# Data Cleaning and Wrangling

```{r, echo=FALSE}
# Columns 20 to 46 have the "%" symbol and I will remove that symbol and convert them to decimals later.
soap <- read_csv("BathSoap.csv", col_types = c("ffffffffffdiidid??d???????????????????????????")) # set column types.
head(soap)
```


```{r, message=FALSE, warning=FALSE}
# Renaming columns so they are easier to work with.
soap %>% 
  rename(id = "Member id", 
         sec = SEC,
         feh = FEH,
         mt = MT,
         sex = SEX,
         age = AGE,
         edu = EDU,
         hs = HS,
         child = CHILD,
         cs = CS,
         affluence = "Affluence Index",
         num_brand = "No. of Brands",
         brand_runs = "Brand Runs",
         tot_vol = "Total Volume",
         num_trans = `No. of  Trans`,
         value = Value,
         trans_ovr_br = "Trans / Brand Runs",
         vol_tran = "Vol/Tran",
         ave_price = "Avg. Price",
         pur_vol_no_promo = "Pur Vol No Promo - %",
         pur_vol_promo = `Pur Vol Promo 6 %`,
         pur_vol_diff_promo = "Pur Vol Other Promo %",
         others_999 = "Others 999",
         br_57_144 = "Br. Cd. 57, 144") %>% 
  rename_at(vars(starts_with("Br. Cd. ")), 
            funs(str_replace(., "Br. Cd. ", "br_"))) %>% 
  rename_at(vars(starts_with("Pr Cat ")), funs(str_replace(., "Pr Cat ", "pr_cat_"))) %>% 
  rename_at(vars(starts_with("PropCat")), funs(str_replace(., "PropCat ", "prop_cat_"))) -> soap
```

```{r}
# Removing all the "%" symbols from columns 18 to 44 and converting them to decimals to represent "percent".
soap[, 20:46] %>% 
  mutate_all(funs(gsub("[[:punct:]]", "", .))) -> soap[, 20:46]

soap[,20:46] <- lapply(soap[,20:46], function(x) as.numeric(x)/100)

# Others 999 brand category is not in percentage but should be. The range is 0 to 10 and the data dictionary says that it should be a percent of volume. A max of 10 doesn't fit that definition. Going to assume that it wasn't converted to a decimal correctly. 
soap$others_999 <- soap$others_999/10
# The sum of the brand categories should total 1 and the mean should be around 1 if I am correct in the others 999 category being incorrectly calculated. 
mean(rowSums(soap[, 23:31]))
# the mean of the row sums for the brand categories is around 1 so the correction I made is the right decision.
```

```{r}
# Checking to see if there are any missing data in our df. 
colMeans(is.na(soap))
```


# Exploratory Data Analysis 

```{r, echo=FALSE}
summary(soap)

soap %>% 
  select(feh, mt, sex, edu, hs, cs) %>% 
  filter(feh == 0 & mt == 0 & sex == 0)
```

Many of the demographics are not specified across many of the same columns and since k-Means uses continuous variables, they are not important to the clustering algorithm but may need to revisit this after clustering to possibly impute values for these 68  customers.

```{r, warning=FALSE, echo=FALSE}
soap[, 2:10] %>% 
  gather(key = Variable, value = Value) %>% 
  ggplot() +
    geom_histogram(aes(x = Value), fill = "steelblue", stat = "count") +
    facet_wrap(~Variable, scales='free') +
    theme_classic() +
    theme(aspect.ratio = 0.5, axis.title = element_blank(), panel.grid = element_blank()) 

soap[, 11:22] %>% 
  gather(key = Variable, value = Value) %>% 
  ggplot() +
    geom_histogram(aes(x = Value), fill = "steelblue") +
    facet_wrap(~Variable, scales='free') +
    theme_classic() +
    theme(aspect.ratio = 0.5, axis.title = element_blank(), panel.grid = element_blank()) 

soap[, 23:31] %>% 
  gather(key = Variable, value = Value) %>% 
  ggplot() +
    geom_histogram(aes(x = Value), fill = "steelblue") +
    facet_wrap(~Variable, scales='free') +
    theme_classic() +
    theme(aspect.ratio = 0.5, axis.title = element_blank(), panel.grid = element_blank()) 

soap[, 32:46] %>% 
  gather(key = Variable, value = Value) %>% 
  ggplot() +
    geom_histogram(aes(x = Value), fill = "steelblue") +
    facet_wrap(~Variable, scales='free') +
    theme_classic() +
    theme(aspect.ratio = 0.5, axis.title = element_blank(), panel.grid = element_blank()) 
```


**Histogram of Variables 2 to 10**
Most of the customers were age level 3 and 4. The child variable has the highest count at level 4 and then level 2. The cs level highest was level 1 and the edu variable has levels 4 and 5 mostly. feh was level 1 and 3. hs was between 3 to 6 and mt was 10. sec was all even in the distribution of counts and sex was 2.

**Histogram of Variables 11 to 22**
Affluence, ave_price, num_trans all seem fairly normally distributed. Affluence has many 0's included that seem that they are missing values so if you impute those values then I suspect the graph will seem more normal. Many of the other variables are skew negative like pur_vol_promo. Purchase volume decreased with promotion but the purchase volume without a promotion increased in volume. 

**Histogram of Variables 23 to 31**
br_57_144 has the most activity out of the brands but it could be that the combination of them caused the increase. Others_999 has a lot of activity but it could be due to representing a lot of different brands. 

**Histogram of Variables 32 to 46**
All other prop_cat seem to have as much activity as 1, 2, and 5. 
prop_cat_1, 2, & 5 need to look at closer.

## Outlier Inspection

```{r}
soap %>% 
  select(feh, mt, sex, edu, hs, cs, affluence) %>%
  filter(affluence == 0)
# Noticed that many of the same columns in the demographics section are also "0" the same as "affluence". I believe that due to the lack of demographic information they were unable to calculate the affluence level. That makes sense.

soap %>% 
  filter(brand_runs > 50) %>% 
  select(affluence, brand_runs, num_trans, vol_tran, ave_price)
# This seems good to me as the affluence level for most is mid-range to high, the number of transactions and total volume is also high so you would expect number of brand runs to be high as well.

soap %>% 
  filter(vol_tran > 1500) %>%
  select(affluence, value, num_trans, tot_vol, vol_tran, ave_price, num_trans) %>%
  mutate(vol_tran_check = tot_vol/num_trans)
# Everything seems okay with these, nothing that seems off or strange like an error in inputting value into df.

soap %>% 
  mutate(vol_tran_check = round(tot_vol/num_trans, 2)) %>% 
  filter(vol_tran == vol_tran_check) %>% 
  select(vol_tran, vol_tran_check)
# checking average volume per transaction to see if there are any calculation errors. 583 of 600 were the same value. Checking the 17 that were different.

soap %>% 
  mutate(vol_tran_check = round(tot_vol/num_trans, 2)) %>% 
  filter(vol_tran != vol_tran_check) %>% 
  select(vol_tran, vol_tran_check) # numbers seem fine, only error due to rounding differences. Nothing significant.
```

## Variable Correlations and Relationship Exploration

**Stronger Positive Correlations**

```{r, echo=FALSE}
cont <- soap[, 11:46]
cc <- cor(soap[, 11:46], method = "pearson")
corrplot(cc, method = "circle", type = "lower")

soap %>% 
  ggplot(aes(color = affluence)) +
    geom_point(aes(x = tot_vol, y = value))

soap %>% 
ggplot(aes(color = affluence)) +
    geom_point(aes(x = pr_cat_1, y = ave_price))

soap %>% 
  ggplot(aes(color = affluence)) +
    geom_point(aes(x = pr_cat_3, y = br_55))

soap %>% 
ggplot(aes(color = affluence)) +
    geom_point(aes(x = prop_cat_11, y = br_481))

soap %>% 
  ggplot() +
    geom_point(aes(x = prop_cat_13, y = br_24), color = "steelblue")

soap %>% 
  ggplot() +
    geom_point(aes(x = prop_cat_14, y = br_55), color = "steelblue")

soap %>% 
  ggplot() +
    geom_point(aes(x = prop_cat_14, y = pr_cat_3), color = "steelblue")
```

**Visualization Summary**
  + value and total volume (the higher the value the higher the volume, typically the larger weight costs more so that makes sense. Since these are different soap products that would make sense.)
  + Pr_cat_1 and ave_price (the higher the average price, the more of product category 1 is purchased. It is probably the price category for more premiumm products.)
  + Pr_cat_3 and br_55 (Brand 55 is in price category 3.)
  + prop_cat_14 and br_55 (I am guessing this means that most likely br_55 is mostly in prop_cat_14, same conclusion for the next two brands below.)
  + prop_cat_11 and br_481
  + prop_cat_13 and br_24
  + prop_cat_3 and pr_cat_14 (proposition category 3 must sell mostly in the third price category.)

## Stronger Negative Correlations Visualizations
```{r, echo=FALSE}
soap %>% 
  ggplot() +
    geom_point(aes(x = pur_vol_no_promo, y = pur_vol_promo), color = "firebrick")

soap %>% 
  ggplot() +
    geom_point(aes(x = pur_vol_diff_promo, y = pur_vol_promo), color = "firebrick")

soap %>% 
  ggplot() +
    geom_point(aes(x = pr_cat_3, y = ave_price), color = "firebrick")

soap %>% 
  ggplot() +
    geom_point(aes(x = prop_cat_14, y = ave_price), color = "firebrick")
```

**Visualization Summary**
  + no promotion and different promotions vs. promotion 6. This makes sense, those who purchase without promotion are making a purchase for a different reason than those who buy during a promotion or purchasing during a different promotion.
  + price category 3 and average price. The lower the average price the more of price cat 3 products. This means that this category has lower priced products.
  + proposition category 14 and average price, this category has lower cost product. brand 55 must also be lower cost on average.

### Exploring Brand Loyalty

```{r, results=FALSE}
library(matrixStats)
# After thinking about how to best represent the brand columns in a way that k-Means will be able to use in the clustering algorithm without the noise of the 0's and very small decimals. I decided to use the max value in each row for the percentage spent. I choose this because it will be a continuous value and also kept the information for measuring brand loyalty: proportion of money spent on a specific brand.  
brand <- as.matrix(soap[23:31])
soap$br_max <- rowMaxs(brand) 

soap %>% 
  relocate(id, sec, feh, mt, sex, age, edu, hs, child, cs, affluence, num_brand, brand_runs, num_trans, tot_vol, value, trans_ovr_br, vol_tran, ave_price, pur_vol_no_promo, pur_vol_promo, pur_vol_diff_promo, br_max) -> soap
soap
```

### Normalization
```{r}
set.seed(15)
norm <- preProcess(soap[12:47], method = c("scale", "center"))
norm <- predict(norm, soap[12:47])
```

# Part 1: Purchasing Behavior
This segmentation is based upon the volume of purchase, the frequency of purchases, average price, the susceptibility to discounts, and the brand loyalty of a household. 

```{r, echo=FALSE}
purchase_behavior <- norm[, c(1:12)] # created a df based on purchase behavior variables.
```

## k Optimization
```{r, echo=FALSE}
fviz_nbclust(purchase_behavior, kmeans, method = "wss")
```

```{r, echo=FALSE}
fviz_nbclust(purchase_behavior, kmeans, method = "silhouette")
```

**4 to 5 clusters would seem to me to be reasonable. Optimal k would be around 4 or 5 due to the "elbow" of the curve being at that point and using the information from the silhouettte method it would be at 5, however, 4 isn't much lower.**

### K-means for k = 4 & 5 Analysis

```{r, echo=FALSE}
purchases_kmeans4 <- kmeans(purchase_behavior, centers = 4, nstart = 25)
purchases_kmeans5 <- kmeans(purchase_behavior, centers = 5, nstart = 25)
```

```{r, echo=FALSE}
fviz_cluster(purchases_kmeans4, data = purchase_behavior) -> purchase_behave
purchase_behave
fviz_cluster(purchases_kmeans5, data = purchase_behavior) -> purchase_behave
purchase_behave
```

```{r, echo=FALSE}
library(cluster)
sil <- silhouette(purchases_kmeans4$cluster, dist(purchase_behavior))
fviz_silhouette(sil)

sil <- silhouette(purchases_kmeans5$cluster, dist(purchase_behavior))
fviz_silhouette(sil)

plot(purchases_kmeans4$centers)
plot(purchases_kmeans5$centers)
```

### **k Selection**
After looking at both k = 4 and k =5, 5 clusters seems to be better as the clusters center distances seem more widely spread out and evenly. Also, they don't seems as close together. It seems to me that the k=5 had less negative values which means they probably put more points into the right clusters. 

## Cluster Analysis
```{r, echo=FALSE, message=FALSE, warning=FALSE}
soap$purch_behavior_cl <- purchases_kmeans5$cluster

norm_soap_centers <- as.data.frame(cbind(c(1, 2, 3, 4, 5), purchases_kmeans5$centers))

soap[, c(15:26,48)]%>%
  group_by(purch_behavior_cl) %>% 
  summarise_all(mean) -> soap_centers

library(reshape2)
cluster_mean <- melt(norm_soap_centers)
cluster_mean$cluster <- c(1,2,3,4,5)
cluster_mean$cluster <- as.factor(cluster_mean$cluster)
cluster_mean <- cluster_mean[-1:-5, ]
cluster_mean$value <- round(cluster_mean$value, 2)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(
  cluster_mean,
  aes(variable, value, group = cluster, color = factor(cluster))) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Variables", y = "Normalized Value", title = "Big spenders, Want Promotions" ) +
  theme(legend.position = "right") +
  theme(axis.text.x = element_text(
    colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 12),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 16)
    ) + 
  scale_color_manual(name = "Cluster", values = c("#FF0000", "#17A589", "#D3D3D3", "#D3D3D3", "#D3D3D3"))
```

### Cluster 1
* The customers in this cluster are likely to be the **big spenders**: 
  + high total volume of transactions 
  + high total value of transactions
 
### Cluster 2
* The customers in this cluster are **value seekers**: 
  + won't make purchases without a promotion
  + will purchased items that have higher average prices
  

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(
  cluster_mean,
  aes(variable, value, group = cluster, color = factor(cluster))
) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Variables", y = "Normalized Value", title = "Brand Loyal, Occasional Purchasers" ) +
  theme(legend.position = "right") +
  theme(axis.text.x = element_text(
    colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 10),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 16)
    ) + 
  scale_color_manual(name = "Cluster", values = c("#D3D3D3", "#D3D3D3", "#FF0000", "#17A589", "#D3D3D3"))
```

### Cluster 3
* The customers in this cluster are **brand loyal**: 
  + most brand loyal as they purchase lowest number of brands
  + spend their money on the brands they are loyal to
  + have a higher number of transactions per brand run
  + are more brand loyal to brands with lower prices 
  + do not need promotions to buy their brands

### Cluster 4
* The customers in this cluster are **occasional purchasers**: 
  + lowest number of transactions
  + lowest volume
  + lowest total value

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(
  cluster_mean,
  aes(variable, value, group = cluster, color = factor(cluster))
) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Variables", y = "Normalized Value", title = "Brand Jumpers" ) +
  theme(legend.position = "right") +
  theme(axis.text.x = element_text(
    colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 10),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 16)
    ) + 
  scale_color_manual(name = "Cluster", values = c("#D3D3D3", "#D3D3D3", "#D3D3D3", "#D3D3D3", "#FF0000"))
```

### Cluster 5
* The customers in this cluster are **not brand loyal**: 
  + Highest number of brands and brand runs
  + highest number of transactions
  + lower transactions per brand
  + lowest percentage spent on a specific brand
  + lower volume per transaction
  + Doesn't matter as much to them to buy when there is a promotion or not
  
  
## Summary
What are the variables that best identify the clusters? Looking at the graphs of each cluster, they all seem important to distinguishing each cluster.
  
  
  
  
  
  
  
  
 
# Part 2: Why Do Customers Buy? The Factors
This segmentation is based average price, price categories, and the selling propositions. 

```{r}
purchase_reason <- norm[, c(8, 22:36)] # created a df based on purchase behavior variables.
```

## k Optimization
```{r, echo=FALSE}
fviz_nbclust(purchase_reason, kmeans, method = "wss")
```

```{r, echo=FALSE}
fviz_nbclust(purchase_reason, kmeans, method = "silhouette")
```

### **k Selection**
3 or 4 clusters would seem to me to be reasonable as the "elbow" of the curve could be interpreted as being at that point. However the silhouettte method is saying at 10 but that is too many clusters and not realistic for this problem. I will start with 3 and 4 and see the results before making a decision.

### K-means for k = 3 & 4 Analysis

```{r,echo=FALSE}
purchases_kmeans3 <- kmeans(purchase_reason, centers = 3, nstart = 25)
purchases_kmeans4 <- kmeans(purchase_reason, centers = 4, nstart = 25)
```

```{r, echo=FALSE}
fviz_cluster(purchases_kmeans4, data = purchase_reason) -> purchase_basis
purchase_basis
fviz_cluster(purchases_kmeans3, data = purchase_reason) -> purchase_basis
purchase_basis
```

```{r, echo=FALSE}
sil <- silhouette(purchases_kmeans4$cluster, dist(purchase_reason))
fviz_silhouette(sil)

sil <- silhouette(purchases_kmeans3$cluster, dist(purchase_reason))
fviz_silhouette(sil)

plot(purchases_kmeans4$centers)
plot(purchases_kmeans3$centers)
```

### **k Selection**
After looking at both k = 3 and k =4, 3 clusters seems to be better as the clusters are less overlapped, clusters are more compact, and centers are spread further apart. 

## Cluster Analysis

```{r, echo=FALSE, warning=FALSE}
soap$purch_reason_cl <- purchases_kmeans3$cluster

norm_soap_centers <- as.data.frame(cbind(c(1, 2, 3), purchases_kmeans3$centers))

soap[, c(8,22:36,49)]%>%
  group_by(purch_reason_cl) %>% 
  summarise_all(mean) -> soap_centers_2

cluster_mean <- melt(norm_soap_centers)
cluster_mean$cluster <- c(1,2,3)
cluster_mean$cluster <- as.factor(cluster_mean$cluster)
cluster_mean <- cluster_mean[-1:-3, ]
cluster_mean$value <- round(cluster_mean$value, 2)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(
  cluster_mean,
  aes(variable, value, group = cluster, color = factor(cluster))) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Variables", y = "Normalized Value", title = "Motivated by Price Category 2 and 4, Prop Category 5") +
  theme(legend.position = "right") +
  theme(axis.text.x = element_text(
    colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 12),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 14)
    ) + 
  scale_color_manual(name = "Cluster", values = c("#FF0000", "#D3D3D3", "#D3D3D3"))
```

### Cluster 1
* The customers in this cluster are likely to make purchases if: 
  + the product's price is in category 2 and 4
  + the product is more in proposition category 5 or but also 11
  + Less likely to purchase a product in price categories 1 and 3 or in proposition category 12 to 14
  

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(
  cluster_mean,
  aes(variable, value, group = cluster, color = factor(cluster))
) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Variables", y = "Normalized Value", title = "Highly motivated by Price Category 1 & Like Most Prop Categories") +
  theme(legend.position = "right") +
  theme(axis.text.x = element_text(
    colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 10),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 12)
    ) + 
  scale_color_manual(name = "Cluster", values = c("#D3D3D3", "#FF0000", "#D3D3D3"))
```

### Cluster 2
* The customers in this cluster are likely to make purchases if: 
  + the average price is higher
  + the product's price is in category 1 (highly motivated to purchase in that category)
  + makes purchases in most of the proposition cateogries
  + Less likely to purchase a product in price categories 2 to 4 or in proposition category 11 and 14
  
  
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(
  cluster_mean,
  aes(variable, value, group = cluster, color = factor(cluster))
) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Variables", y = "Normalized Value", title = "Motivated by Price Category 2 and 4, Prop Category 5" ) +
  theme(legend.position = "right") +
  theme(axis.text.x = element_text(
    colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 10),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 12)
    ) + 
  scale_color_manual(name = "Cluster", values = c("#D3D3D3", "#D3D3D3", "#FF0000"))
```

### Cluster 3
* The customers in this cluster are likely to make purchases if: 
  + the average price very low
  + the product's price is in category 2 and 4
  + Motivated to purchase in prop_cat_5

## Summary
* What are the variables that identify the clusters:
  + Average Price
  + All Price Categories
  + All proposition categories except maybe numbers 9, 11, and 15 as most of the values for each cluster is closer to 0, the center. I am interpreting those as not distinguishable factors between clusters.






  
  
## Part 3: Purchase Behavior and Basis of Purchase
How can we use the above knowledge to create clusters that combine what we have learned so far? I am going to take the variables from both parts where at least one cluster's mean was at least 1 SD from the center, 0. The variables I will be using are all of the variables for part 1 and all the variables from part 2. I will see if removing prop_cat_9, 11, and 15 has any impact on the clusters by keeping them in and then removing them to see if there is any impact on the clusters.

```{r}
purch_behavior_basis <- norm[, c(1:12, 22:29,31,33:35)] # created a df based on purchase behavior and basis variables.
```

## k Optimization
```{r, echo=FALSE}
fviz_nbclust(purch_behavior_basis, kmeans, method = "wss")
```

```{r, echo=FALSE}
fviz_nbclust(purch_behavior_basis, kmeans, method = "silhouette")
```

**3 to 4 clusters would seem to me to be reasonable due to the previous sections and the above measures help to confirm to start our investigation with those values for k. Also, I decided to remove prop_cat_9, 11, and 15 after running the model with them. The clusters were overlapping a lot more and were the distances of the centers were very close together.**

### K-means for k = 3 & 4 Analysis

```{r, echo=FALSE}
purchases_kmeans3 <- kmeans(purch_behavior_basis, centers = 3, nstart = 25)
purchases_kmeans4 <- kmeans(purch_behavior_basis, centers = 4, nstart = 25)
```

```{r, echo=FALSE}
fviz_cluster(purchases_kmeans4, data = purch_behavior_basis) -> purchase_behave
purchase_behave
fviz_cluster(purchases_kmeans3, data = purch_behavior_basis) -> purchase_behave
purchase_behave
```

```{r, echo=FALSE}
library(cluster)
sil <- silhouette(purchases_kmeans3$cluster, dist(purch_behavior_basis))
fviz_silhouette(sil)

sil <- silhouette(purchases_kmeans4$cluster, dist(purch_behavior_basis))
fviz_silhouette(sil)
```

## **k Selection**
After looking at both k = 3 and k =4, 3 clusters seems to be better as the clusters are more defined and there is less overlap. Also from our previous sections, three clusters would be more useful for classifying the types of customers in each.

## Cluster Analysis
```{r, echo=FALSE, warning=FALSE}
soap$purch_behavior_basis_cl <- purchases_kmeans3$cluster

norm_soap_centers <- as.data.frame(cbind(c(1, 2, 3), purchases_kmeans3$centers))

soap[, c(12:23, 33:40, 42, 44:46, 50)]%>%
  group_by(purch_behavior_basis_cl) %>% 
  summarise_all(mean) -> soap_centers

cluster_mean <- melt(norm_soap_centers)
cluster_mean$cluster <- c(1,2,3)
cluster_mean$cluster <- as.factor(cluster_mean$cluster)
cluster_mean <- cluster_mean[-1:-3, ]
cluster_mean$value <- round(cluster_mean$value, 2)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(
  cluster_mean,
  aes(variable, value, group = cluster, color = factor(cluster))) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Variables", y = "Normalized Value", title = "Brand Loyal, Price Category 3, and Prop Category 14" ) +
  theme(legend.position = "right") +
  theme(axis.text.x = element_text(
    colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 12),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 12)
    ) + 
  scale_color_manual(name = "Cluster", values = c("#FF0000", "#D3D3D3", "#D3D3D3"))

soap %>% 
  filter(purch_behavior_basis_cl == 1) %>%
  ggplot() +
    geom_histogram(aes(x = affluence), fill = "steelblue")

soap[, c(2:10, 50)] %>% 
  filter(purch_behavior_basis_cl == 1) %>% 
  gather(key = Variable, value = Value) %>% 
  ggplot() +
    geom_histogram(aes(x = Value), fill = "steelblue", stat = "count") +
    facet_wrap(~Variable, scales='free') +
    theme_classic() +
    theme(aspect.ratio = 0.5, axis.title = element_blank(), panel.grid = element_blank()) 
```

### **Cluster 1: Purchase Behavior and Basis**
* The customers in this cluster are likely to be the **brand loyal** customers who: 
  + purchase items of a lower price
  + respond to products in the third price category and proposition category 14
  
### **Cluster 1 Demographics** 
* The customers in this cluster are:
  + lower affluence
  + Women of ages category 2 to 4
  + non-vegetarian
  + Have child categories 2, 4, and 5
  + mostly have native language of 10
  + have 4 and 5 members in the household
  + come from socioeconomic lower status of 4
  + mostly, lower amounts of education 
  
 
```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(
  cluster_mean,
  aes(variable, value, group = cluster, color = factor(cluster))
) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Variables", y = "Normalized Value", title = "Brand Loyal, Occasional Purchasers" ) +
  theme(legend.position = "right") +
  theme(axis.text.x = element_text(
    colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 10),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 12)
    ) + 
  scale_color_manual(name = "Cluster", values = c("#D3D3D3", "#FF0000", "#D3D3D3"))

soap %>%
  filter(purch_behavior_basis_cl == 2) %>%
  ggplot() +
    geom_histogram(aes(x = affluence), fill = "steelblue")

soap[, c(2:10, 50)] %>% 
  filter(purch_behavior_basis_cl == 2) %>% 
  gather(key = Variable, value = Value) %>% 
  ggplot() +
    geom_histogram(aes(x = Value), fill = "steelblue", stat = "count") +
    facet_wrap(~Variable, scales='free') +
    theme_classic() +
    theme(aspect.ratio = 0.5, axis.title = element_blank(), panel.grid = element_blank()) 
```

### **Cluster 2: Purchase Behavior and Basis**
* The customers in this cluster move around brands and: 
  + base their purchases on the price of the product
  + they are not motivated by the promotions
  + their purchases are mostly in price categories 2 and 4 and prop category 5
  
### **Demographics** 
* The customers in this cluster are:
  + mostly lower to higher middle affluence
  + Women of ages category 3 to 4
  + Have child categories 2 and 4
  + education of 4 to 5
  + non-vegetarian
  + mostly have native language of 4 and 10
  + have 3 to 6 members in the household
  + come from socioeconomic lower status of 3 and 4


```{r, echo=FALSE, message=FALSE, warning=FALSE}
ggplot(
  cluster_mean,
  aes(variable, value, group = cluster, color = factor(cluster))
) +
  geom_line() +
  scale_color_viridis_d() +
  labs(x = "Variables", y = "Normalized Value", title = "Premium Customers Who Love a Good Deal" ) +
  theme(legend.position = "right") +
  theme(axis.text.x = element_text(
    colour = "grey20", size = 12, angle = 90, hjust = 0.5, vjust = 0.5),
                        axis.text.y = element_text(colour = "grey20", size = 10),
                        strip.text = element_text(face = "italic"),
                        text = element_text(size = 12)
    ) + 
  scale_color_manual(name = "Cluster", values = c("#D3D3D3", "#D3D3D3", "#FF0000"))

soap %>%
  filter(purch_behavior_basis_cl == 3) %>%
  ggplot() +
    geom_histogram(aes(x = affluence), fill = "steelblue")

soap[, c(2:10, 50)] %>% 
  filter(purch_behavior_basis_cl == 3) %>% 
  gather(key = Variable, value = Value) %>% 
  ggplot() +
    geom_histogram(aes(x = Value), fill = "steelblue", stat = "count") +
    facet_wrap(~Variable, scales='free') +
    theme_classic() +
    theme(aspect.ratio = 0.5, axis.title = element_blank(), panel.grid = element_blank()) 

```

### **Cluster 3: Purchase Behavior and Basis**
* The customers in this cluster are **premium customers**: 
  + they purchase products that have a high average price
  + they could be more brand loyal
  + they love a good promotion
  + they love the products in price category 1 and any proposition category that is 6 to 13
  
### **Demographics** 
* The customers in this cluster are:
  + higher affluence
  + Women of ages category 3 to 4
  + vegetarian and non-vegetarian
  + Have child category of 4
  + mostly have native language of 4 and 10
  + have 4 and 5 members in the household
  + come from socioeconomic lower status of 3 and 4 
  + higher education levels
  







## Part 4: Predictive Model
We need to create a model that will predict if a customer will be in cluster 1.


```{r, warning=FALSE}
library(fastDummies)
soap[, 51:54] <- dummy_cols(soap$purch_behavior_basis_cl)

# Removing unnecessary variables to make coding easier in this section
soap2 <- soap[, c(-1, -24:-32, -41, -43, -47:-51)] 

soap2$.data_1 <- as.factor(soap2$.data_1)


# split into soap dataset into training and validation set to test the predictive power of the model.
p <- createDataPartition(soap2$.data_1,p=0.7,list=FALSE)
train <- as.data.frame(soap2[p,])
valid <- as.data.frame(soap2[-p,])

# Applying logistic regression model 
model1 <- glm(formula = .data_1 ~ cs+edu+feh+hs+sex+brand_runs+trans_ovr_br+value+ave_price+pr_cat_2+pr_cat_3+prop_cat_5+prop_cat_14+br_max, family = binomial, 
    data = train[, -36:-37])
# choose variables to include in logistic regression model that showed to be predictive of cluster 1 customers based on the discoveries made in part 3 cluster 1 discussion. Also, using logistic regression due to the categorical and continuous variables.

predict_validation<-predict(model1, newdata = valid, type='response')

library(e1071)

## Categorizing the result based on the cutoff value(0.5)
resultcheck<-as.factor(ifelse(predict_validation>0.5,1,0))
confusionMatrix(resultcheck, valid$.data_1)

# The model has a specificity of 95%, correctly identifying customers who should be in cluster 1.

```