---
title: "mbruner3_1.r"
author: "Mark Bruner"
date: "9/11/2020"
output: html_document
---
## Reference
Data is taken from http://faculty.marshall.usc.edu/gareth-james/ISL/College.csv

## Start
Loaded in the college.csv and assigned it to the variable "college". Renamed the first column in the data set.
```{r}
library(readr)
college <- read_csv("college.csv", col_types = "cciiiiiiiiiiiiiiiii")
college <- rename(college, "College Name" = "X1")
college 
```

Loaded "summarytools" package.
```{r}
library(summarytools)
```

# Quantitative Descriptive Statistics
Made the "Apps" column in college data set a variable and ran summary of desc. statistcs.
```{r}
apps <- college$Apps
descr(x = apps)
```

##Qualitative Descriptive Statistics & Transformation of Data 
Used the "Private" column in the College data set and converted it to a DF. After that I added a column to the DF to represent the percentage.
```{r}
private <- table(college$Private)
private_df <- as.data.frame(private)
private_df <- rename(private_df, "Private" = "Var1")
private_percent <- private_df$Freq/sum(private_df$Freq)
private_df_percent <- add_column(private_df, private_percent)
private_df_percent <- rename(private_df_percent, "Percent" = "private_percent")
private_df_percent
```
## Scatterplot: Multi-variable
Created a scatterplot of the two variable Applications and Acceptance in the College Data Set.
```{r}
library(ggplot2)
Applications = college$Apps
Acceptance = college$Accept
Private = college$Private
ggplot(data = college, aes(x = Applications, y = Acceptance, col = Private, shape = Private)) + geom_point()
```

# Single Variable Visualization
ggplot(data = college, aes(x = Private)) + geom_histogram(stat = "Count", fill = "light blue", col = "skyblue")

```{r}
ggplot(data = college, aes(x = Private)) + geom_histogram(stat = "Count", fill = "light blue", col = "skyblue")
```