---
title: "mbruner3_3"
author: "Mark Bruner"
date: "10/4/2020"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

# Libraries needed for this assignment.
```{r, echo=FALSE}
library(tidyverse)
library(lubridate)
library(caret)
library(ISLR)
library(fastDummies)
library(dplyr)
library(naivebayes)
```
# Imported dataset.
```{r, echo=FALSE}
flight_delays <- read_csv("FlightDelays.csv", col_types = "?f?ficfffffff")
flight_delays$FL_DATE <- mdy(flight_delays$FL_DATE) # converted the FL_DATE column to date format.
head(flight_delays)
```
# Getting to know data.
```{r}
summary(flight_delays)
```

```{r}
str(flight_delays)
```
**The five main carriers for the flights are DH, RU, US, DL, and MQ. Over half of the flights are to LGA (La Guardia, NY). The mean and median a fairly close together for the distance variable which means that the distribution is symmetrical. Most of the flights originated out of DCA (Washington DC). The weather only impacted 32/2201 flights and caused delays from January 25 to January 27th mostly. More flights occurred on days 4 and 5. Lastly, about 19% of all flights were delayed in January.**

# Checked for missing values and created dummy variables.
```{r}
colMeans(is.na(flight_delays)) # no missing data.
```
**Since all variables have column means of 0, that implies that there are no missing values in this dataset.**

# Created "bins" for CRS_DEP_TIME (Scheduled Departure Time).
```{r}
library(OneR)
flight_delays_time <- as.data.frame(bin(flight_delays$CRS_DEP_TIME, nbins = 18, labels = c(1:18)))

flight_delays <- cbind(flight_delays_time, flight_delays)

flight_delays <- flight_delays %>% rename("CRS_DEP_TIME GROUP NO" = "bin(flight_delays$CRS_DEP_TIME, nbins = 18, labels = c(1:18))") 

flight_delays %>%
  group_by(`CRS_DEP_TIME GROUP NO`) %>%
  summarise(CRS_DEP_TIME = length(`CRS_DEP_TIME GROUP NO`))

flight_delays <- flight_delays[, -2]
head(flight_delays)
```

# Separated dataset into only predictors and then partitioned it into training and vaildation sets.
```{r}
flight_delays_predictors <- flight_delays[, c(1, 3, 4, 7, 10, 13)] # created df with predictors.

set.seed(15)
index_train <- createDataPartition(flight_delays_predictors$DAY_WEEK, p = .6, list = FALSE)
flight_train <- flight_delays_predictors[index_train, ] 
flight_valid <- flight_delays_predictors[-index_train, ]
head(flight_train)
```
# Naive Bayes Model with training data.
```{r}
naive_model <- naive_bayes(flight_train[, 1:5], flight_train[, 6], laplace = 1)
```

# Predicting delayed/on-time flights on validation dataset.
```{r}
predicted_flight_labels <- predict(naive_model, flight_valid, type = "class")
```

# Destination and Origin Proportion and Count Tables
```{r}
table(predicted_flight_labels, flight_valid$`Flight Status` , dnn = c("Prediction", "Actual"))
dest_prop <- as.data.frame(naive_model$tables$DEST)
dest_count_delay <- as.data.frame(dest_prop[1:3, 3]*832)
dest_count_delay <- rename(dest_count_delay, Count = `dest_prop[1:3, 3] * 832`)
dest_count_ontime <- as.data.frame(dest_prop[4:6, 3]*46)
dest_count_ontime <- rename(dest_count_ontime, Count = `dest_prop[4:6, 3] * 46`)

dest_count <- rbind(dest_count_delay, dest_count_ontime)
dest_count_prop <- cbind(dest_prop, dest_count)

origin_prop <- as.data.frame(naive_model$tables$ORIGIN)
origin_count_delay <- as.data.frame(origin_prop[1:3, 3]*832)
origin_count_delay <- rename(origin_count_delay, Count = `origin_prop[1:3, 3] * 832`)
origin_count_ontime <- as.data.frame(origin_prop[4:6, 3]*46)
origin_count_ontime <- rename(origin_count_ontime, Count = `origin_prop[4:6, 3] * 46`)

origin_count <- rbind(origin_count_delay, origin_count_ontime)
origin_count_prop <- cbind(origin_prop, origin_count)

dest_count_prop <- rename(dest_count_prop, "Ontime/Delayed" = Var2, Proportion = "Freq")
dest_count_prop

origin_count_prop <- rename(origin_count_prop, "Ontime/Delayed" = Var2, Proportion = "Freq")
origin_count_prop
```
**This models predicts that JFK will have 8 delays, LGA will have 20 delays, and that EWR will have 18 delays. Also, that BWI will have 4 delays, DCA will have 23 delays, and that IAD will have 18 delays.**

```{r}
library(gmodels)
confusionMatrix(flight_valid$Delayed, predicted_flight_labels, dnn = c("Actual", "Prediction"))
```

**This model needs refining since incorrectly predicted that 148 flights would be delayed but were actually on-time giving it an "okay" sensitivity of 80% but still needs optimization. More importantly, the specificity (predicting a flight will be delayed and is actually delayed) is 46% which is not good since we want this model to predict if a flight will be delayed.** 


```{r}
library(pROC)
predicted_flight_labels <- predict(naive_model, flight_valid, type = "prob")
roc(flight_valid$Delayed, predicted_flight_labels[, 2])

plot.roc(flight_valid$Delayed, predicted_flight_labels[, 2])
```
**The area under this curve is 64.6% which tells us that this model can be further improved since we want area to be reasonably close to 100%.** 



