---
title: "mbruner3_3"
author: "Mark Bruner"
date: "10/4/2020"
output: pdf_document
---

# Libraries needed for this assignment.
```{r}
library(tidyverse)
library(lubridate)
library(caret)
library(ISLR)
library(e1071)
library(fastDummies)
library(dplyr)
```
# Imported dataset.
```{r}
flight_delays <- read_csv("FlightDelays.csv", col_types = "?f?ficfffffff")
flight_delays$FL_DATE <- mdy(flight_delays$FL_DATE) # converted the FL_DATE column to date format.
```

# Cleaned dataset and created dummy variables.
```{r}
colMeans(is.na(flight_delays)) # no missing data.

flight_delays <- dummy_cols(flight_delays, select_columns = "Flight Status", remove_selected_columns = TRUE)
flight_delays <- flight_delays %>% rename("On Time" = "Flight Status_ontime") 
flight_delays <- flight_delays %>% rename("Delayed" = "Flight Status_delayed") 
flight_delays$`On Time` <- as.factor(flight_delays$`On Time`)
flight_delays$Delayed <- as.factor(flight_delays$Delayed)
```
# Created "bins" for CRS_DEP_TIME.
```{r}
library(OneR)
flight_delays_time <- as.data.frame(bin(flight_delays$CRS_DEP_TIME, nbins = 18, labels = c(1:18)))
flight_delays <- cbind(flight_delays_time, flight_delays)
flight_delays <- flight_delays %>% rename("CRS_DEP_TIME GROUP NO" = "bin(flight_delays$CRS_DEP_TIME, nbins = 18, labels = c(1:18))") 
flight_delays %>%
  group_by(`CRS_DEP_TIME GROUP NO`) %>%
  summarise(CRS_DEP_TIME = length(`CRS_DEP_TIME GROUP NO`))
flight_delays <- flight_delays[, -2]
head(flight_delays)
```

# Created "bins" for DEP_TIME.
```{r}
flight_delays_dep_time <- as.data.frame(bin(flight_delays$DEP_TIME, nbins = 18, labels = c(1:18)))
flight_delays <- cbind(flight_delays_dep_time, flight_delays)
flight_delays <- flight_delays %>% rename("DEP_TIME GROUP NO" = "bin(flight_delays$DEP_TIME, nbins = 18, labels = c(1:18))") 
flight_delays %>% 
  group_by(`DEP_TIME GROUP NO`) %>%
  summarise(DEP_TIME = length(`DEP_TIME GROUP NO`))
flight_delays <- flight_delays[, -4]
head(flight_delays)
```
# separated dataset into only predictors and partitioned into training and vaildation set.
```{r}
flight_delays_predictors <- flight_delays[, c(2, 3, 4, 8, 10, 14)] # created df with predictors.

set.seed(15)
index_train <- createDataPartition(flight_delays_predictors$DAY_WEEK, p = .6, list = FALSE)
flight_train <- flight_delays_predictors[index_train, ] 
flight_valid <- flight_delays_predictors[-index_train, ]
head(flight_train)
```

```{r}
naive_model <- naiveBayes(Delayed~`CRS_DEP_TIME GROUP NO` + CARRIER + DEST + ORIGIN + DAY_WEEK, data = flight_train)
dest_prop_table <- as.data.frame.table(naive_model$tables$DEST)
origin_prop_table <- as.data.frame.table(naive_model$tables$ORIGIN)

dest_counts_table <- as.data.frame.table((naive_model$tables$DEST)*2201)
origin_counts_table <- as.data.frame.table((naive_model$tables$ORIGIN)*2201)

dest_table <- cbind(dest_counts_table, dest_prop_table[, 3])
dest_table <- dest_table %>% rename("On-Time or Delayed" = "Y", "Counts" = "Freq", "Proportion" = `dest_prop_table[, 3]`) 
dest_table

origin_table <- cbind(origin_counts_table, origin_prop_table[, 3])
origin_table <- origin_table %>% rename("On-Time or Delayed" = "Y", "Counts" = "Freq", "Proportion" = `origin_prop_table[, 3]`) 
origin_table
```
This model is predicting that JFK will have 17% of all delayed flights between the three airports. LGA will have 44% of all delayed flights and EWR will have 39% of all delayed flights.

```{r}
predicted_flight_labels <- predict(naive_model, flight_valid)

library(gmodels)
confusionMatrix(flight_valid$Delayed, predicted_flight_labels)
```
This model needs refining since incorrectly predicted that 147 flights would be delayed but were actually on-time. 

```{r}
predicted_flight_labels <- predict(naive_model, flight_valid, type = "raw")
head(predicted_flight_labels)
```
```{r}
library(pROC)
roc(flight_valid$Delayed, predicted_flight_labels[, 2])

plot.roc(flight_valid$Delayed, predicted_flight_labels[, 2])
```




